### 2016-12-13 paper/5_evaluation.tex
Insert paragraphs and remove some text to fit it on 20 pages

\begin{figure}[htbp]
	\centering
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{loglogaxis}[
				xlabel={Dimensionality $n$},
				ylabel={Execution time $t$ in s},
				width=\dualpgfwidth,
				legend entries={Local, Spark, Flink},
				legend pos=north west,
			]
			\addplot[
				color=blue,
				mark=x,
			] table[
				x=RowsA,
				y=Time,
			]
			{data/matrixMult/matrixMultLoadReference};
			\addplot[
				color=red,
				mark=o,
			] table[
				x=RowsA,
				y=Time,
			]
			{data/matrixMult/matrixMultLoadSpark};
			\addplot[
				color=teal,
				mark=triangle,
			] table[
				x=RowsA,
				y=Time,
			]
			{data/matrixMult/matrixMultLoadFlink};
			\end{loglogaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:mmLoadRuntime}
	\end{subfigure}
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{semilogxaxis}[
				xlabel={Number of cores},
				ylabel={Execution time $t$ in s},
				legend pos=north west,
				legend entries={Spark, Flink},
				ymin=0.0,
				ymax=2000,
				width=\dualpgfwidth,
			]
			\addplot[
				color=red,
				mark=o,
			] table[
				x=Parallelism,
				y=Time,
			]
			{data/matrixMult/matrixMultScalabilitySpark};
			\addplot[
				color=teal,
				mark=triangle,
			] table[
				x=Parallelism,
				y=Time,
			]
			{data/matrixMult/matrixMultScalabilityFlink};
			\end{semilogxaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:mmNodesRuntime}
	\end{subfigure}
	\caption{Scalability of matrix multiplication. \subref{fig:mmLoadRuntime} Execution time of matrix multiplication depending on the data size. \subref{fig:mmNodesRuntime} Execution time of matrix multiplication depending on the cluster size with constant work load per core.}
	\label{fig:mmScalability}
\end{figure}

@@

\begin{figure}[htbp]
	\centering
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{loglogaxis}[
				xlabel={Number of rows of matrix $V$},
				ylabel={Execution time $t$ in s},
				width=\dualpgfwidth,
				legend pos=north west,
				legend columns=2,
				legend entries={Local, Spark, Flink, HT Spark, HT Flink},
				ymax=500,
			]
			\addplot[
				color=blue,
				mark=x,
			] table[
				x=Rows,
				y=Time,
			]
			{data/nnmfStep/nnmfStepLoadReference};

@@

			\addplot[
				color=teal,
				mark=triangle,
			]table[
				x=Rows,
				y=Time,
			]
			{data/nnmfStep/nnmfStepLoadFlink};
			\addplot[
				color=black,
				mark=diamond,
			]table[
				x=rowsV,
				y=time,
			]
			{data/nnmfStep/gnmfStepBenchLoadSpark};
			\addplot[
				color=magenta,
				mark=square,
			]table[
				x=rowsV,
				y=time,
			]
			{data/nnmfStep/gnmfStepBenchLoadFlink};
			\end{loglogaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:nmfLoadRuntime}
	\end{subfigure}
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{semilogxaxis}[
				xlabel={Number of rows of matrix $V$},
				ylabel={Speedup of HT over Gilbert},
				width=\dualpgfwidth,
				ymin=0,
				ymax=2.5,
				legend entries={Spark,Flink},
				legend pos=north west,
			]
			\addplot[
				color=red,
				mark=o,
			] table[
				x=Rows,
				y=Speedup,
			]
			{data/nnmfStep/gnmfStepSpeedupSpark};

			\addplot[
				color=teal,
				mark=diamond,
			] table[
				x=Rows,
				y=Speedup,
			]
			{data/nnmfStep/gnmfStepSpeedupFlink};
			\end{semilogxaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:gnmfSpeedup}
	\end{subfigure}
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{semilogxaxis}[
				xlabel={Number of cores},
				ylabel={Execution time $t$ in s},
				xmax=101,
				ymin=0,
				ymax=60,
				legend entries={Spark, Flink},
				legend pos=north west,
				width=\dualpgfwidth,
			]
			\addplot[
				color=red,
				mark=o,
			] table[
				x=Parallelism,
				y=Time,
			]
			{data/nnmfStep/nnmfStepScalabilitySpark};

@@

% \subsection{Optimizations}

% In this section, we want to evaluate the effects Gilbert's optimizer has on the runtime of programs.
% For this purpose, we execute one GNMF step on Gilbert's Stratosphere and Spark execution engine using the same settings as in \cref{subsec:NMF}.
% However, this time we disable the transpose pushdown and matrix-multiplication reordering optimization.
% The results are shown in \cref{fig:nnmfLoadOptimization}.

% \begin{figure}
% 	\centering
% 	\begin{subfigure}[t]{\dualpgfwidth}
% 		\begin{tikzpicture}
% 			\begin{semilogxaxis}[
% 				xlabel={Rows $d$ of $V$},
% 				ylabel={Execution time $t$ in s},
% 				legend pos=north west,
% 				legend entries={Optimized Spark, Optimized Stratosphere, Non-optimized Spark, Non-optimized Stratosphere},
% 				width=\dualpgfwidth,
% 			]

% 			\addplot[
% 				color=red,
% 				mark=o,
% 			] table[
% 				x=Rows,
% 				y=Time,
% 			]
% 			{data/nnmfStepLoadNoOp/nnmfStepLoadSpark};

% 			\addplot[
% 				color=teal,
% 				mark=triangle,
% 			]table[
% 				x=Rows,
% 				y=Time,
% 			]
% 			{data/nnmfStepLoadNoOp/nnmfStepLoadStratosphere};

% 			\addplot[
% 				color=blue,
% 				mark=x,
% 			] table[
% 				x=Rows,
% 				y=Time,
% 			]
% 			{data/nnmfStepLoadNoOp/nnmfStepNoOpLoadSpark};

% 			\addplot[
% 				color=black,
% 				mark=diamond,
% 			]table[
% 				x=Rows,
% 				y=Time,
% 			]
% 			{data/nnmfStepLoadNoOp/nnmfStepNoOpLoadStratosphere};

% 			\end{semilogxaxis}
% 		\end{tikzpicture}
% 		\caption{}
% 		\label{fig:nnmfLoadOptimization}	
% 	\end{subfigure}
% 	\begin{subfigure}[t]{\dualpgfwidth}
% 		\begin{tikzpicture}
% 			\begin{semilogxaxis}[
% 				xlabel={Rows $d$ of $V$},
% 				ylabel={Speedup},
% 				legend pos=north west,
% 				legend entries={Optimized Spark, Optimized Stratosphere},
% 				width=\dualpgfwidth,
% 			]

% 			\addplot[
% 				color=red,
% 				mark=o,
% 			] table[
% 				x=Rows,
% 				y=Speedup,
% 			]
% 			{data/nnmfStepLoadNoOp/gnmfStepLoadSpeedupSpark};

% 			\addplot[
% 				color=teal,
% 				mark=triangle,
% 			]table[
% 				x=Rows,
% 				y=Speedup,
% 			]
% 			{data/nnmfStepLoadNoOp/gnmfStepLoadSpeedupStratosphere};			
% 			\end{semilogxaxis}
% 		\end{tikzpicture}
% 		\caption{}
% 		\label{fig:gnmfLoadOptimizationSpeedup}
% 	\end{subfigure}
% 	\caption{Effect of Gilbert's optimizations using the example of a single GNMF step executed on a $50$-core cluster. \subref{fig:nnmfLoadOptimization} Execution time of the optimized and non-optimized GNMF step depending on the input size. \subref{fig:gnmfLoadOptimizationSpeedup} Speedup of the optimized GNMF step with respect to the non-optimized GNMF step depending on the input size.}
% \end{figure}

% We can observe that the Gilbert optimizer has a significant effect on the runtime of GNMF.
% The optimized Gilbert program runs on both distributed engines faster than it's non-optimized counterparts.
% Interestingly, the Spark execution engine seems to benefit more from the optimizer than the Stratosphere executor.
% The speedups of the optimized GNMF step with respect to its non-optimized counterparts are shown in \cref{fig:gnmfLoadOptimizationSpeedup}
% For the optimized program running on Spark's execution engine, a steadily ascending speedup, except for $d=500$, can be observed.
% The optimized Spark version runs about $18$ times faster than the non-optimized program for $d=10000$.
% In contrast to that, Stratosphere shows a constant speedup of roughly $1.7$ between the optimized and non-optimized version.

% It is not yet known what causes these different speedups.
% However, looking at the GNMF formula in \cref{lst:nmf}, it becomes obvious why the optimization works.
% There are two matrix multiplications that can be optimized: $W^T WH$ in line $6$ and $WHH^T$ in line $7$.
% Keeping in mind that $W\in \mathbb{R}^{d\times 10}$ and $H\in \mathbb{R}^{10 \times 100000}$, we can assess the different strategies.
% The non-optimized version would execute the matrix multiplications due to left-associativity from left to right:

% \begin{displaymath}
% 	\overbrace{\underbrace{\left(W^T W\right)}_{\in \mathbb{R}^{10\times 10}}H}^{\in\mathbb{R}^{10\times 100000}}
% \end{displaymath}

% and

% \begin{displaymath}
% 	\overbrace{\underbrace{\left(WH\right)}_{\in \mathbb{R}^{d \times 100000}}H^T}^{\in \mathbb{R}^{d\times 10}}
% \end{displaymath}

% For the first matrix multiplication, the execution order is optimal, since the intermediate result of $W^T W$ is much smaller than $WH$.
% However, for the second matrix multiplication, the execution order is far from optimal.
% The left-associativity produces an intermediate matrix result of size $d\times 100000$.
% For large $d$ this matrix becomes really huge.
% Additionally, the intermediate result is dense, because the operands $W$ and $H$ are dense, as well.
% By changing the execution order, we can decrease the size of the intermediate result significantly.

% \begin{displaymath}
% 	\overbrace{W\underbrace{\left(HH^T\right)}_{\in \mathbb{R}^{10 \times 10}}}^{\in \mathbb{R}^{d\times 10}}
% \end{displaymath}

% The Gilbert optimizer first detects the matrix multiplication reordering and then changes it so that the maximum intermediate result is minimized.
% That is the reason why the optimized GNMF program runs clearly faster.

@@

\begin{listing}[htbp]
	\begin{lstlisting}[language=Matlab,
		commentstyle=\color{black},
		  stringstyle=\color{black},
		  keywordstyle=\color{black}\bfseries,
		  morekeywords={ones, fixpoint},
		  xleftmargin=.0\textwidth]
% load adjacency matrix
A = load();
maxIterations = 10;
d = sum(A, 2); % outdegree per vertex
% create the column-stochastic transition matrix
T = (diag(1 ./ d) * A)'; 
r_0 = ones(numVertices, 1) / numVertices; % initialize the ranks
e = ones(numVertices, 1) / numVertices;
% PageRank calculation
fixpoint(r_0, @(r) .85 * T * r + .15 * e, maxIterations)
	\end{lstlisting}
	\caption{Gilbert PageRank implementation.}
	\label{lst:gilbertPageRank}
\end{listing}

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.3\linewidth]{images/pageRankStep.png}
	\caption{Data flow of one iteration of the PageRank algorithm for Spark and Flink.}
	\label{fig:pageRankDataFlow}
\end{figure}

@@

\begin{figure}[htbp]
	\centering
	\begin{subfigure}[h]{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{loglogaxis}[
				xlabel={Number of vertices $n$},
				ylabel={Execution time $t$ in s},
				xmax=250000,
				ymax=50000,
				legend pos=north west,
				legend entries={Spark, Flink, SP Flink, SP Spark},
				width=\dualpgfwidth,
			]

@@

			\addplot[black,
				mark=triangle,
			] table[
				x=rows,
				y=time,
			]
			{data/pagerank/pagerankBenchLoadSpark};
			\end{loglogaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:pageRankResults}
	\end{subfigure}
	\begin{subfigure}[h]{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{semilogxaxis}[
				xlabel={Number of vertices $n$},
				ylabel={Speedup},
				legend pos=north west,
				legend entries={SP Spark, SP Flink},
				width=\dualpgfwidth,
			]

>>>>

\begin{figure}[htbp]
	\centering
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{loglogaxis}[
				xlabel={Dimensionality $n$},
				ylabel={Execution time $t$ in s},
				width=\dualpgfwidth,
				height=.78\dualpgfwidth,
				legend entries={Local, Spark, Flink},
				legend pos=north west,
				font=\footnotesize
			]
			\addplot[
				color=blue,
				mark=x,
			] table[
				x=RowsA,
				y=Time,
			]
			{data/matrixMult/matrixMultLoadReference};
			\addplot[
				color=red,
				mark=o,
			] table[
				x=RowsA,
				y=Time,
			]
			{data/matrixMult/matrixMultLoadSpark};
			\addplot[
				color=teal,
				mark=triangle,
			] table[
				x=RowsA,
				y=Time,
			]
			{data/matrixMult/matrixMultLoadFlink};
			\end{loglogaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:mmLoadRuntime}
	\end{subfigure}
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{semilogxaxis}[
				xlabel={Number of cores},
				ylabel={Execution time $t$ in s},
				legend pos=north west,
				legend entries={Spark, Flink},
				ymin=0.0,
				ymax=2000,
				width=\dualpgfwidth,
				height=.78\dualpgfwidth,
				font=\footnotesize
			]
			\addplot[
				color=red,
				mark=o,
			] table[
				x=Parallelism,
				y=Time,
			]
			{data/matrixMult/matrixMultScalabilitySpark};
			\addplot[
				color=teal,
				mark=triangle,
			] table[
				x=Parallelism,
				y=Time,
			]
			{data/matrixMult/matrixMultScalabilityFlink};
			\end{semilogxaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:mmNodesRuntime}
	\end{subfigure}
	\caption{Scalability of matrix multiplication. \subref{fig:mmLoadRuntime} Execution time of matrix multiplication depending on the data size. \subref{fig:mmNodesRuntime} Execution time of matrix multiplication depending on the cluster size with constant work load per core.}
	\label{fig:mmScalability}
\end{figure}

@@

\begin{figure}[htbp]
	\centering
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{loglogaxis}[
				xlabel={Number of rows of matrix $V$},
				ylabel={Execution time $t$ in s},
				width=\dualpgfwidth,
				height=.78\dualpgfwidth,
				legend pos=north west,
				legend columns=2,
				legend entries={Local, Spark, Flink, HT Spark, HT Flink},
				ymax=800,
				font=\footnotesize
			]
			\addplot[
				color=blue,
				mark=x,
			] table[
				x=Rows,
				y=Time,
			]
			{data/nnmfStep/nnmfStepLoadReference};

@@

			\addplot[
				color=teal,
				mark=triangle,
			]table[
				x=Rows,
				y=Time,
			]
			{data/nnmfStep/nnmfStepLoadFlink};
			\addplot[
				color=black,
				mark=diamond,
			]table[
				x=rowsV,
				y=time,
			]
			{data/nnmfStep/gnmfStepBenchLoadSpark};
			\addplot[
				color=magenta,
				mark=square,
			]table[
				x=rowsV,
				y=time,
			]
			{data/nnmfStep/gnmfStepBenchLoadFlink};
			\end{loglogaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:nmfLoadRuntime}
	\end{subfigure}
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{semilogxaxis}[
				xlabel={Number of rows of matrix $V$},
				ylabel={Speedup of HT over Gilbert},
				width=\dualpgfwidth,
				height=.78\dualpgfwidth,
				ymin=0,
				ymax=2.5,
				legend entries={Spark,Flink},
				legend pos=north west,
				font=\footnotesize
			]
			\addplot[
				color=red,
				mark=o,
			] table[
				x=Rows,
				y=Speedup,
			]
			{data/nnmfStep/gnmfStepSpeedupSpark};

			\addplot[
				color=teal,
				mark=diamond,
			] table[
				x=Rows,
				y=Speedup,
			]
			{data/nnmfStep/gnmfStepSpeedupFlink};
			\end{semilogxaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:gnmfSpeedup}
	\end{subfigure}
	\begin{subfigure}{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{semilogxaxis}[
				xlabel={Number of cores},
				ylabel={Execution time $t$ in s},
				xmax=101,
				ymin=0,
				ymax=60,
				legend entries={Spark, Flink},
				legend pos=north west,
				width=\dualpgfwidth,
				height=.7\dualpgfwidth,
				font=\footnotesize
			]
			\addplot[
				color=red,
				mark=o,
			] table[
				x=Parallelism,
				y=Time,
			]
			{data/nnmfStep/nnmfStepScalabilitySpark};

@@

@@

\begin{listing}[htbp]
	\begin{lstlisting}[language=Matlab,
		commentstyle=\color{black},
		  stringstyle=\color{black},
		  keywordstyle=\color{black}\bfseries,
		  morekeywords={ones, fixpoint},
		  xleftmargin=.0\textwidth,
		  basicstyle=\footnotesize]
% load adjacency matrix
A = load();
maxIterations = 10;
d = sum(A, 2); % outdegree per vertex
% create the column-stochastic transition matrix
T = (diag(1 ./ d) * A)'; 
r_0 = ones(numVertices, 1) / numVertices; % initialize the ranks
e = ones(numVertices, 1) / numVertices;
% PageRank calculation
fixpoint(r_0, @(r) .85 * T * r + .15 * e, maxIterations)
	\end{lstlisting}
	\caption{Gilbert PageRank implementation.}
	\label{lst:gilbertPageRank}
\end{listing}

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.275\linewidth]{images/pageRankStep.png}
	\caption{Data flow of one iteration of the PageRank algorithm for Spark and Flink.}
	\label{fig:pageRankDataFlow}
\end{figure}

@@

\begin{figure}[htbp]
	\centering
	\begin{subfigure}[h]{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{loglogaxis}[
				xlabel={Number of vertices $n$},
				ylabel={Execution time $t$ in s},
				xmax=250000,
				ymax=50000,
				legend pos=north west,
				legend entries={Spark, Flink, SP Flink, SP Spark},
				width=\dualpgfwidth,
				height=.78\dualpgfwidth,
				font=\footnotesize
			]

@@

			\addplot[black,
				mark=triangle,
			] table[
				x=rows,
				y=time,
			]
			{data/pagerank/pagerankBenchLoadSpark};
			\end{loglogaxis}
		\end{tikzpicture}
		\caption{}
		\label{fig:pageRankResults}
	\end{subfigure}
	\begin{subfigure}[h]{\dualpgfwidth}
		\begin{tikzpicture}
			\begin{semilogxaxis}[
				xlabel={Number of vertices $n$},
				ylabel={Speedup},
				legend pos=north west,
				legend entries={SP Spark, SP Flink},
				width=\dualpgfwidth,
				height=.78\dualpgfwidth,
				font=\footnotesize
			]
